Visual Interrogation of Attention-Based Models for Natural Language Inference and Machine Comprehension

### 2018-08-17 EMNLP-demo-paper/result.tex
fix quote mark problem

As illustrated in Fig.~\ref{fig:NLIexample}(a), the input sentence pair (P:``A woman in a green jacket is drinking tea.'' H:``A woman is drinking green tea.'') is predicted to be \emph{entailment}, which is incorrect.
By examining the attention, we can see the word \emph{green} in ``\emph{green} jacket" is aligned to the \emph{green} in ``\emph{green} tea". However, these two \emph{greens} modify different nouns, which potentially leads to the wrong prediction. The grammatical structure is visually shown in the form of the dependency tree.
%i.e., \emph{green} in \textbf{P} is attached to ``jacket''.
However, the model does not have access to the syntactic information and mistakenly assumes the two \emph{greens} modify the same thing.
% (thus predict \emph{entailment}). 
%
To correct the mistake, we can edit the attention value and remove the align between these two "greens" (see (c)(b)).
As expected, the prediction label is corrected (\emph{neutral}).

@@

As illustrated in Figure~\ref{fig:MCexample}, we represent \emph{att2} as colored bars with a \emph{yellow-green-blue} colormap. Each rectangular bar corresponds to one sentence. The user can focus on individual sentences by clicking on the rectangular bar (see Figure~\ref{fig:MCexample}).
%The proposed visualization help reveal the potential alignment issues in the machine comprehension model.
The $p1$, $p2$ colored bars (\emph{white-red} colormap) illustrate the predicted probabilities of the start and the end index for the answer (the deeper the higher). %The span of text defined by the beginning and end index with the highest probability is the answer.
%
%In Figure~\ref{fig:MCexample} ($a3$), there are two spans of text that correspond to higher probabilities. 
In Figure~\ref{fig:MCexample}(a3), the sentence containing the answer exhibits good alignment with the question (e.g., ``Two'' with ``many'').
%In (a3), we examine the sentence that contains the answer, which exhibits excellent alignment with the question (e.g., ``Two'' is aligned with ``many''). 
%The spans of text with high probability are ``Two'' and ``No'' (highlight in orange), both align well with the word ``many''.
Interestingly, the number "9" (in "Falcon 9") is also aligned with "many", which may lead to problems.

The user can explore the robustness of the model by examining how the prediction varies when the question is perturbed.
%
As illustrated in ($b1$, and $b2$), the perturbation removes the word "partial" in the original sentence, which leads the model to produce a different yet correct answer ("No"). Referring to ($a1$), we can see the word "No" exhibits the second highest probability for the original question.
%
The user can also manually edit the text. As shown in ($b3$), the model still produces the correct answer when changing the question from "how many" to "which".
%One possible explanation is that the word ``Two'' has a much higher \emph{att2} attention value compared to ``9'', which may contribute to the current outcome.

>>>>

As illustrated in Fig.~\ref{fig:NLIexample}(a), the input sentence pair (P:``A woman in a green jacket is drinking tea.'' H:``A woman is drinking green tea.'') is predicted to be \emph{entailment}, which is incorrect.
By examining the attention, we can see the word \emph{green} in ``\emph{green} jacket" is aligned to the \emph{green} in ``\emph{green} tea". However, these two \emph{greens} modify different nouns, which potentially leads to the wrong prediction. The grammatical structure is visually shown in the form of the dependency tree.
%i.e., \emph{green} in \textbf{P} is attached to ``jacket''.
However, the model does not have access to the syntactic information and mistakenly assumes the two \emph{greens} modify the same thing.
% (thus predict \emph{entailment}). 
%
To correct the mistake, we can edit the attention value and remove the align between these two ``greens" (see (c)(b)).
As expected, the prediction label is corrected (\emph{neutral}).

@@

As illustrated in Figure~\ref{fig:MCexample}, we represent \emph{att2} as colored bars with a \emph{yellow-green-blue} colormap. Each rectangular bar corresponds to one sentence. The user can focus on individual sentences by clicking on the rectangular bar (see Figure~\ref{fig:MCexample}).
%The proposed visualization help reveal the potential alignment issues in the machine comprehension model.
The $p1$, $p2$ colored bars (\emph{white-red} colormap) illustrate the predicted probabilities of the start and the end index for the answer (the deeper the higher). %The span of text defined by the beginning and end index with the highest probability is the answer.
%
%In Figure~\ref{fig:MCexample} ($a3$), there are two spans of text that correspond to higher probabilities. 
In Figure~\ref{fig:MCexample}(a3), the sentence containing the answer exhibits good alignment with the question (e.g., ``Two'' with ``many'').
%In (a3), we examine the sentence that contains the answer, which exhibits excellent alignment with the question (e.g., ``Two'' is aligned with ``many''). 
%The spans of text with high probability are ``Two'' and ``No'' (highlight in orange), both align well with the word ``many''.
Interestingly, the number ``9" (in ``Falcon 9") is also aligned with ``many", which may lead to problems.

The user can explore the robustness of the model by examining how the prediction varies when the question is perturbed.
%
As illustrated in ($b1$, and $b2$), the perturbation removes the word ``partial" in the original sentence, which leads the model to produce a different yet correct answer (``No"). Referring to ($a1$), we can see the word ``No" exhibits the second highest probability for the original question.
%
The user can also manually edit the text. As shown in ($b3$), the model still produces the correct answer when changing the question from ``how many" to ``which".
%One possible explanation is that the word ``Two'' has a much higher \emph{att2} attention value compared to ``9'', which may contribute to the current outcome.
